### 2023-08-23 sections/2-methods.tex
added quicksort partition and better detail in sieve algorithm

\subsection {Index-Building}
\begin{itemize}
    \item sharded search 
    \item index building 
    \item optimization of number of shards 
\end{itemize}

@@

We start by letting $Q$ be a vector of clusters and points sorted by non-decreasing $\delta_{max}$ and initializing $Q$ with the root cluster. 
While we haven't found a threshold which guarantees exactly $k$ points, we repeat the following:

\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item For each cluster left in $Q$, if it is a leaf or has $k$ or fewer points, remove it and all of its points to $Q$.
\item Replace all remaining non-leaf clusters in $Q$ with their children. 
\end{enumerate}

The points remaining in $Q$ are the $k$ nearest neighbors.

TODO: Add find kth algorithm.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:sieve1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[5] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \WHILE{$|Q| > k$}
            \STATE $C_{\tau} \leftarrow Q.findKth.\delta_{max}$
            \FOR {$q \in Q$}
                \IF {$q.\delta_{min} > \tau$}
                    \STATE $Q.remove(q)$
                \ENDIF
            \ENDFOR
            \FOR {$q \in Q$}
                \IF {$q.isLeaf \lor |q.cardinality| \leq k$}
                    \STATE $Q.push(q.points)$
                \ELSE
                    \STATE $[l, r] \leftarrow q.children$
                    \STATE $Q.push([l, r])$   
                \ENDIF
                \STATE $Q.remove(q)$
            \ENDFOR 
        \ENDWHILE
        \STATE Return $Q$
    \end{algorithmic}
    \end{algorithm}

>>>>

\subsection {Index-Building}
\begin{itemize}
    \item sharded search
    \item dataset depth-first reordering 
    \item index building 
    \item optimization of number of shards 
\end{itemize}

@@

We start by letting $Q$ be a vector initialized with the root cluster. This vector will maintain the 
clusters and points which are still in contention for being one of the nearest neighbors.
While $Q$ contains at least one cluster (i.e., $Q$ is not a vector of only points), we repeat the process described in 
the following paragraphs.

First, 
we aim to find the element $q_{\tau} \in Q$ with the smallest $\delta_{max}$ such that 
$q_{\tau}$ and all elements in $Q$ with smaller $\delta_{max}$ collectively contain at least $k$ points.

To find this element, 
we use a process similar to the Partition algorithm used in Quicksort (CITE: original Quicksort paper and reference wherever it is 
in our paper), adjusted to account for the varying cardinalities of elements in $Q$. Our modified algorithm finds the smallest index $i$ such 
that all elements in $Q$ with $\delta_{max}$ closer to or equal to $Q[i]$'s $\delta_{max}$ collectively have cardinality greater
than or equal to $k$. In the process of finding this index, we alter the order of $Q$ so that all elements to left
of index $i$ have $\delta_{max}$ less than or equal to $Q[i]$'s $\delta_{max}$ and all elements to right of index $i$
have $\delta_{max}$ greater than $Q[i]$'s $\delta_{max}$. We consider $Q[i]$ to be our $q_{\tau}$, and our threshold 
$\tau$ to be $Q[i]$'s $\delta_{max}$.

We then remove from $Q$ any element whose $\delta_{min}$ is greater than $\tau$. Then, for each element left 
in $Q$, if it is a leaf or contains $k$ or fewer points, we remove it and all of its points to $Q$. Finally, we
replace all remaining non-leaf clusters in $Q$ with their children.

We complete this process until $Q$ contains only points. At this point, we 
just use [INSERT ref to quicksort partition in paper] to find the $k$th 
nearest neighbor. Since this algorithm will also elements on the correct side of 
the index of the $k$th nearest neighbor, the first $k$ elements in $Q$ a are the $k$-nearest neighbors.

\begin{algorithm} % enter the algorithm environment
    \caption{QuicksortPartition(\emph{Q, k, l, r})} % give the algorithm a caption
    \label{alg:quicksort-partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[5] % enter the algorithmic environment
        \IF{$l \geq r$}
            \STATE Return $min(l, r)$
        \ELSE 
            \STATE $pivot = l + (r - l) / 2$
            \STATE $Q.swap(pivot, r)$
            \STATE $a \leftarrow l$
            \STATE $b \leftarrow l$
            \WHILE {$b < r$}
                \IF {$Q[b].\delta_{max} < Q[r].\delta_{max}$}
                    \STATE $Q.swap(a, b)$
                    \STATE $a \leftarrow a + 1$
                \ENDIF
                \STATE $b \leftarrow b + 1$
            \ENDWHILE
            \STATE $Q.swap(a, r)$
            \STATE $p \leftarrow r$
            \STATE $g \leftarrow \sum_{i=0}^{p} Q[i].cardinality$
            \IF {$g = k$}
                \STATE Return $p$
            \ELSIF {$g < k$}
                \STATE $QuicksortPartition(Q, k, p + 1, r)$
            \ELSE 
                \IF {$p > 0 \land g > k + Q[p - 1].cardinality$}
                    \STATE $QuicksortPartition(Q, k, l, p - 1)$
                \ELSIF {$p > 0 \land g = k + Q[p - 1].cardinality$}
                    \STATE Return $p - 1$
                \ELSE
                    \STATE Return $p$
                \ENDIF
            \ENDIF
        \ENDIF
    \end{algorithmic}
    \end{algorithm}

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve(\emph{tree, query, k})} % give the algorithm a caption
    \label{alg:sieve} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[6] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ [$tree.root$]
        \WHILE{$|Q| > k$}
            \STATE $i \leftarrow QuicksortPartition(Q, k, 0, |Q| - 1)$
            \STATE $\tau \leftarrow Q[i].\delta_{max}$
            \FOR {$q \in Q$}
                \IF {$q.\delta_{min} > \tau$}
                    \STATE $Q.remove(q)$
                \ENDIF
            \ENDFOR
            \FOR {$q \in Q$}
                \IF {$q.isLeaf \lor |q.cardinality| \leq k$}
                    \STATE $Q.push(q.points)$
                \ELSE
                    \STATE $[l, r] \leftarrow q.children$
                    \STATE $Q.push([l, r])$   
                \ENDIF
                \STATE $Q.remove(q)$
            \ENDFOR 
        \ENDWHILE
        \STATE Return $Q$
    \end{algorithmic}
    \end{algorithm}
